---
title: "DATA363_ResearchProject"
author: "Micaila Marcelle"
date: "2025-04-09"
output: pdf_document
---

# TODO box plot for monetary

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE TOTAL DISTANCE BAR CHARTS
################################################################################

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
races <- read_csv('https://docs.google.com/spreadsheets/d/1j5ezobC_HXnCS8S18cZg6zQaJqisjrUvqVK8b8x3v08/gviz/tq?tqx=out:csv')
races <- races[1:13]
races
```

First, we need to clean up the data after reading it in. Some of the numeric columns were read as strings, so those columns need to be fixed.
```{r}
races$`Individual Winner Monetary Prize - MEN` <- gsub("^€", "",  races$`Individual Winner Monetary Prize - MEN`)
races$`Total Prize Pool - MEN` <- gsub("^€", "",  races$`Total Prize Pool - MEN`)
races$`Individual Winner Monetary Prize - WOMEN` <- gsub("^€", "",  races$`Individual Winner Monetary Prize - WOMEN`)
races$`Total Prize Pool - WOMEN` <- gsub("^€", "",  races$`Total Prize Pool - WOMEN`)
races$`Total Distance - MEN`<- gsub("km", "", races$`Total Distance - MEN`)
races$`Total Distance - WOMEN`<- gsub("km", "", races$`Total Distance - WOMEN`)
races$`Total Distance - MEN` <- as.double(races$`Total Distance - MEN`)
races$`Total Distance - WOMEN` <- as.double(races$`Total Distance - WOMEN`)
races$`Individual Winner Monetary Prize - MEN` <- as.numeric(gsub(',', '', races$`Individual Winner Monetary Prize - MEN`))
races$`Individual Winner Monetary Prize - WOMEN` <- as.numeric(gsub(',', '', races$`Individual Winner Monetary Prize - WOMEN`))
races$`Total Prize Pool - MEN` <- as.numeric(gsub(',', '', races$`Total Prize Pool - MEN`))
races$`Total Prize Pool - WOMEN` <- as.numeric(gsub(',', '', races$`Total Prize Pool - WOMEN`))
races$`Final Winner Time - MEN` <- as.numeric(races$`Final Winner Time - MEN`)/60
races$`Final Winner Time - WOMEN` <- as.numeric(races$`Final Winner Time - WOMEN`)/60
```


```{r}
summary(races)
```

First, the mean total distance for each year for the women's races is calculated, with these values being entered in manually for now. Note that all of these values are in kilometers, and that the races are separated by category. 
```{r}
wMeanDist_df = data.frame(Year=c(2016:2024))
wMeanDist_df$DistStage = c(0)
wMeanDist_df$DistGT = c(0)
wMeanDist_df$DistOD = c(0)

for(i in 0:8) {
  wMeanDist_df$DistStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Total Distance - WOMEN`)
  wMeanDist_df$DistGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Total Distance - WOMEN`)
  if(i >= 7) {
    wMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Total Distance - WOMEN`)
  } else {
    wMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Total Distance - WOMEN`)
  }
}
```


The same means are then calculated for the men's races:
```{r}
mMeanDist_df = data.frame(Year=c(2016:2024))
mMeanDist_df$DistStage = c(0)
mMeanDist_df$DistGT = c(0)
mMeanDist_df$DistOD = c(0)

for(i in 0:8) {
  mMeanDist_df$DistStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Total Distance - MEN`)
  mMeanDist_df$DistGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Total Distance - MEN`)
  mMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Total Distance - MEN`)
}
```

We then use this information to construct a sequence of bar charts:
```{r}
distances_matrix <- matrix(c(wMeanDist_df$DistOD, mMeanDist_df$DistOD), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's One-Day Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))

distances_matrix <- matrix(c(wMeanDist_df$DistStage, mMeanDist_df$DistStage), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's Stage Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


distances_matrix <- matrix(c(wMeanDist_df$DistGT, mMeanDist_df$DistGT), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's Grand Tour Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
scatter_plot1 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="One-Day Races Total Distances")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```

```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

scatter_plot2 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="Stage Races Total Distances")
legend("topleft", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

scatter_plot3 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="Grand Tour Races Total Distances")
legend("right", legend=c('Men', 'Women'), pch = 19, col=colors)
```


################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE NUMBER OF TEAMS BAR CHARTS
################################################################################

First, the mean number of teams for each year for the women's races is calculated, with the data values being entered manually, and with the data being split between the different possible categories of races
```{r}
wMeanTeams_df = data.frame(Year=c(2016:2024))
wMeanTeams_df$TeamStage = c(0)
wMeanTeams_df$TeamGT = c(0)
wMeanTeams_df$TeamOD = c(0)

for(i in 0:8) {
  wMeanTeams_df$TeamStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Num Teams - WOMEN`)
  wMeanTeams_df$TeamGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Num Teams - WOMEN`)
  if(i >= 7) {
    wMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Num Teams - WOMEN`)
  } else {
    wMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Num Teams - WOMEN`)
  }
}
```

Next, we find the same values for the men's races:
```{r}
mMeanTeams_df = data.frame(Year=c(2016:2024))
mMeanTeams_df$TeamStage = c(0)
mMeanTeams_df$TeamGT = c(0)
mMeanTeams_df$TeamOD = c(0)

for(i in 0:8) {
  mMeanTeams_df$TeamStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Num Teams - MEN`)
  mMeanTeams_df$TeamGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Num Teams - MEN`)
  mMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Num Teams - MEN`)
}
```

We then use this information to construct a sequence of bar charts:
```{r}
teams_matrix <- matrix(c(wMeanTeams_df$TeamOD, mMeanTeams_df$TeamOD), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's One-Day Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


teams_matrix <- matrix(c(wMeanTeams_df$TeamStage, mMeanTeams_df$TeamStage), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's Stage Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


teams_matrix <- matrix(c(wMeanTeams_df$TeamGT, mMeanTeams_df$TeamGT), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's Grand Tour Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="One-Day Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="Stage Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="Grand Tour Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```

################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE PLOTS FOR FINAL WINNER TIMES
################################################################################

```{r}
wMeanTime_df = data.frame(Year=c(2016:2024))
wMeanTime_df$TimeStage = c(0)
wMeanTime_df$TimeGT = c(0)
wMeanTime_df$TimeOD = c(0)

for(i in 0:8) {
  wMeanTime_df$TimeStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Final Winner Time - WOMEN`)
  wMeanTime_df$TimeGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Final Winner Time - WOMEN`)
  if(i >= 7) {
    wMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Final Winner Time - WOMEN`)
  } else {
    wMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Final Winner Time - WOMEN`)
  }
}
```

```{r}
mMeanTime_df = data.frame(Year=c(2016:2024))
mMeanTime_df$TimeStage = c(0)
mMeanTime_df$TimeGT = c(0)
mMeanTime_df$TimeOD = c(0)

for(i in 0:8) {
  mMeanTime_df$TimeStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Final Winner Time - MEN`)
  mMeanTime_df$TimeGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Final Winner Time - MEN`)
  mMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Final Winner Time - MEN`)
}
```

```{r}
time_matrix <- matrix(c(wMeanTime_df$TimeOD, mMeanTime_df$TimeOD), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's One-Day Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


time_matrix <- matrix(c(wMeanTime_df$TimeStage, mMeanTime_df$TimeStage), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's Stage Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


time_matrix <- matrix(c(wMeanTime_df$TimeGT, mMeanTime_df$TimeGT), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's Grand Tour Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="One-Day Races Final Winner Times")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="Stage Races Final Winner Times")
legend("topleft", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="Grand Tour Races Final Winner Times")
legend("right", legend=c('Men', 'Women'), pch = 19, col=colors)
```


################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE BOX PLOTS FOR THE INDIVIDUAL AND 
OVERALL PRIZE MONEY
################################################################################

First, we collect the data for the individual and overall prize money for the women's races, with these being separated by category:

```{r}
# I'm thinking 6 box plots/ 3 twin box plots where we show a male and female for stage,GT,OD
# do this both for individual prize and total prize
prize_df <- races[races$Year == 2024,2:6]

womPrizeTotalOD <- prize_df[prize_df$Type == 'One-Day' | prize_df$Type == "Grand Tour / women's one-day" | prize_df$Type == "Men's stage, women's one-day",]$`Total Prize Pool - WOMEN`
womPrizeIndOD <- prize_df[prize_df$Type == 'One-Day' | prize_df$Type == "Grand Tour / women's one-day" | prize_df$Type == "Men's stage, women's one-day",]$`Individual Winner Monetary Prize - WOMEN`

womPrizeTotalStage <- prize_df[prize_df$Type == 'Stage',]$`Total Prize Pool - WOMEN`
womPrizeIndStage <- prize_df[prize_df$Type == 'Stage',]$`Individual Winner Monetary Prize - WOMEN`

womPrizeTotalGT <- prize_df[prize_df$Type == 'Grand Tour',]$`Total Prize Pool - WOMEN`
womPrizeIndGT <- prize_df[prize_df$Type == 'Grand Tour',]$`Individual Winner Monetary Prize - WOMEN`

menPrizeTotalOD <- prize_df[prize_df$Type == 'One-Day',]$`Total Prize Pool - MEN`
menPrizeIndOD <- prize_df[prize_df$Type == 'One-Day',]$`Individual Winner Monetary Prize - MEN`

menPrizeTotalStage <- prize_df[prize_df$Type == 'Stage' | prize_df$Type == "Men's stage, women's one-day",]$`Total Prize Pool - MEN`
menPrizeIndStage <- prize_df[prize_df$Type == 'Stage' | prize_df$Type == "Men's stage, women's one-day",]$`Individual Winner Monetary Prize - MEN`

menPrizeTotalGT <- prize_df[prize_df$Type == 'Grand Tour' | prize_df$Type == "Grand Tour / women's one-day",]$`Total Prize Pool - MEN`
menPrizeIndGT <- prize_df[prize_df$Type == 'Grand Tour' | prize_df$Type == "Grand Tour / women's one-day",]$`Individual Winner Monetary Prize - MEN`
```


###
Twin Total Prize bar plots
###

```{r}
boxplot(c(min(min(menPrizeTotalOD, na.rm=TRUE), min(womPrizeTotalOD, na.rm=TRUE)), max(max(menPrizeTotalOD, na.rm=TRUE), max(womPrizeTotalOD, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeTotalOD, add=TRUE, at=-0.25)
boxplot(menPrizeTotalOD, add=TRUE, at=0.25)


boxplot(c(min(min(menPrizeTotalStage, na.rm=TRUE), min(womPrizeTotalStage, na.rm=TRUE)), max(max(menPrizeTotalStage, na.rm=TRUE), max(womPrizeTotalStage, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeTotalStage, add=TRUE, at=-0.25)
boxplot(menPrizeTotalStage, add=TRUE, at=0.25)


boxplot(c(min(min(menPrizeTotalGT, na.rm=TRUE), min(womPrizeTotalGT, na.rm=TRUE)), max(max(menPrizeTotalGT, na.rm=TRUE), max(womPrizeTotalGT, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeTotalGT, add=TRUE, at=-0.25)
boxplot(menPrizeTotalGT, add=TRUE, at=0.25)
```




```{r}
boxplot(c(min(min(menPrizeIndOD, na.rm=TRUE), min(womPrizeIndOD, na.rm=TRUE)), max(max(menPrizeIndOD, na.rm=TRUE), max(womPrizeIndOD, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeIndOD, add=TRUE, at=-0.25)
boxplot(menPrizeIndOD, add=TRUE, at=0.25)


boxplot(c(min(min(menPrizeIndStage, na.rm=TRUE), min(womPrizeIndStage, na.rm=TRUE)), max(max(menPrizeIndStage, na.rm=TRUE), max(womPrizeIndStage, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeIndStage, add=TRUE, at=-0.25)
boxplot(menPrizeIndStage, add=TRUE, at=0.25)


boxplot(c(min(min(menPrizeIndGT, na.rm=TRUE), min(womPrizeIndGT, na.rm=TRUE)), max(max(menPrizeIndGT, na.rm=TRUE), max(womPrizeIndGT, na.rm=TRUE))), boxfill=NA, border=NA, at=0)
boxplot(womPrizeIndGT, add=TRUE, at=-0.25)
boxplot(menPrizeIndGT, add=TRUE, at=0.25)
```

####
# Total Prize plots Individually
###

```{r}
boxplot(womPrizeTotalOD)
boxplot(womPrizeTotalStage)
boxplot(womPrizeTotalGT)
```


```{r}
boxplot(menPrizeTotalOD)
boxplot(menPrizeTotalStage)
boxplot(menPrizeTotalGT)
```

###
# Individual Prize plots
###

```{r}
boxplot(womPrizeIndOD)
boxplot(womPrizeIndStage)
boxplot(womPrizeIndGT)
```


```{r}
boxplot(menPrizeIndOD)
boxplot(menPrizeIndStage)
boxplot(menPrizeIndGT)
```

################################################################################
 FOLLOWING CODE IS USED FOR GENERATING THE BOX PLOTS FOR THE INDIVIDUAL AND 
 OVERALL PRIZE MONEY
 ################################################################################
 
 First, we collect the data for the individual and overall prize money for the women's races, with these being separated by category:
```{r}
 individualWomenPrizeMoney2024Stage <- c(12000, 12030, 16000)
 individualWomenPrizeMoney2024GrandTour <- c(50000, 50000)
 individualWomenPrizeMoney2024OneDay <- c(16000, 16000, 2256, 16000, 20000, 21000, 16000, 10000, 12500)
 
 overallWomenPrizeMoney2024Stage <- c(100000, 100000, 130000)
 overallWomenPrizeMoney2024GrandTour <- c(250000, 250000)
 overallWomenPrizeMoney2024OneDay <- c(43000, 40000, 40000, 50000, 34000, 20000)
```
 
 
 We then collect the same data for the men's races:
```{r}
 individualMenPrizeMoney2024Stage <- c(12000, 35500, 12000, 16000, 12300, 20000)
 individualMenPrizeMoney2024GrandTour <- c(150000, 265000, 500000)
 individualMenPrizeMoney2024OneDay <- c(16000, 16000, 16000, 16000, 16000, 20000, 30000, 16000, 16000, 20000)
 
 overallMenPrizeMoney2024Stage <- c(100000, 300000, 99000, 130000, 103700, 128000)
 overallMenPrizeMoney2024GrandTour <- c(1119960, 1600000, 2573202)
 overallMenPrizeMoney2024OneDay <- c(43000, 40000, 40000, 40000, 40000, 50000, 91000, 34000, 40000, 50000)
```

A barplot is then constructed for the individual and overall in each category:
```{r}
ind_mean_tableOD <- c(mean(individualWomenPrizeMoney2024OneDay), mean(individualMenPrizeMoney2024OneDay))
barplot(ind_mean_tableOD, xlab="", ylab="Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024OneDay) + 5000), col=c("lightblue", "lightcoral"), main="One Day Races Individual Winner Mean Prize Money (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

ind_mean_tableS <- c(mean(individualWomenPrizeMoney2024Stage), mean(individualMenPrizeMoney2024Stage))
barplot(ind_mean_tableS, xlab="", ylab="Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024Stage) + 5000), col=c("lightblue", "lightcoral"), main="Stage Races Individual Winner Mean Prize Money (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

ind_mean_tableGT <- c(mean(individualWomenPrizeMoney2024GrandTour), mean(individualMenPrizeMoney2024GrandTour))
barplot(ind_mean_tableGT, xlab="", ylab="Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024GrandTour) + 5000), col=c("lightblue", "lightcoral"), main="Grand Tour Races Individual Winner Mean Prize Money (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableOD <- c(mean(overallWomenPrizeMoney2024OneDay), mean(overallMenPrizeMoney2024OneDay))
barplot(overall_mean_tableOD, xlab="", ylab="Overall Mean Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024OneDay) + 5000), col=c("lightblue", "lightcoral"), main="One Day Races Overall Mean Prize Pool (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableS <- c(mean(overallWomenPrizeMoney2024Stage), mean(overallMenPrizeMoney2024Stage))
barplot(overall_mean_tableS, xlab="", ylab="Overall Mean Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024Stage) + 5000), col=c("lightblue", "lightcoral"), main="Stage Races Overall Mean Prize Pool (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableGT <- c(mean(overallWomenPrizeMoney2024GrandTour), mean(overallMenPrizeMoney2024GrandTour))
barplot(overall_mean_tableGT, xlab="", ylab="Overall Mean Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024GrandTour) + 5000), col=c("lightblue", "lightcoral"), main="Grand Tour Races Overall Mean Prize Pool (Euros)", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))
```




