---
title: "DATA363_ResearchProject"
author: "Micaila Marcelle"
date: "2025-04-09"
output: pdf_document
---

# TODO box plot for monetary

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE TOTAL DISTANCE BAR CHARTS
################################################################################

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
races <- read_csv('https://docs.google.com/spreadsheets/d/1j5ezobC_HXnCS8S18cZg6zQaJqisjrUvqVK8b8x3v08/gviz/tq?tqx=out:csv')
races <- races[1:13]
races
```

First, we need to clean up the data after reading it in. Some of the numeric columns were read as strings, so those columns need to be fixed.
```{r}
races$`Individual Winner Monetary Prize - MEN` <- gsub("^€", "",  races$`Individual Winner Monetary Prize - MEN`)
races$`Total Prize Pool - MEN` <- gsub("^€", "",  races$`Total Prize Pool - MEN`)
races$`Individual Winner Monetary Prize - WOMEN` <- gsub("^€", "",  races$`Individual Winner Monetary Prize - WOMEN`)
races$`Total Prize Pool - WOMEN` <- gsub("^€", "",  races$`Total Prize Pool - WOMEN`)
races$`Total Distance - MEN`<- gsub("km", "", races$`Total Distance - MEN`)
races$`Total Distance - WOMEN`<- gsub("km", "", races$`Total Distance - WOMEN`)
races$`Total Distance - MEN` <- as.double(races$`Total Distance - MEN`)
races$`Total Distance - WOMEN` <- as.double(races$`Total Distance - WOMEN`)
races$`Individual Winner Monetary Prize - MEN` <- as.numeric(gsub(',', '', races$`Individual Winner Monetary Prize - MEN`))
races$`Individual Winner Monetary Prize - WOMEN` <- as.numeric(gsub(',', '', races$`Individual Winner Monetary Prize - WOMEN`))
races$`Total Prize Pool - MEN` <- as.numeric(gsub(',', '', races$`Total Prize Pool - MEN`))
races$`Total Prize Pool - WOMEN` <- as.numeric(gsub(',', '', races$`Total Prize Pool - WOMEN`))
races$`Final Winner Time - MEN` <- as.numeric(races$`Final Winner Time - MEN`)/60
races$`Final Winner Time - WOMEN` <- as.numeric(races$`Final Winner Time - WOMEN`)/60
```


```{r}
summary(races)
```

First, the mean total distance for each year for the women's races is calculated, with these values being entered in manually for now. Note that all of these values are in kilometers, and that the races are separated by category. 
```{r}
wMeanDist_df = data.frame(Year=c(2016:2024))
wMeanDist_df$DistStage = c(0)
wMeanDist_df$DistGT = c(0)
wMeanDist_df$DistOD = c(0)

for(i in 0:8) {
  wMeanDist_df$DistStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Total Distance - WOMEN`)
  wMeanDist_df$DistGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Total Distance - WOMEN`)
  if(i >= 7) {
    wMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Total Distance - WOMEN`)
  } else {
    wMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Total Distance - WOMEN`)
  }
}
```


The same means are then calculated for the men's races:
```{r}
mMeanDist_df = data.frame(Year=c(2016:2024))
mMeanDist_df$DistStage = c(0)
mMeanDist_df$DistGT = c(0)
mMeanDist_df$DistOD = c(0)

for(i in 0:8) {
  mMeanDist_df$DistStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Total Distance - MEN`)
  mMeanDist_df$DistGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Total Distance - MEN`)
  mMeanDist_df$DistOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Total Distance - MEN`)
}
```

We then use this information to construct a sequence of bar charts:
```{r}
distances_matrix <- matrix(c(wMeanDist_df$DistOD, mMeanDist_df$DistOD), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's One-Day Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))

distances_matrix <- matrix(c(wMeanDist_df$DistStage, mMeanDist_df$DistStage), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's Stage Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


distances_matrix <- matrix(c(wMeanDist_df$DistGT, mMeanDist_df$DistGT), nrow = 2, byrow = TRUE)
rownames(distances_matrix) <- c("Womens", "Mens")
colnames(distances_matrix) <- wMeanDist_df$Year
barplot(distances_matrix, xlab="Year", ylab="Mean Distance (km)", main="Men and Women's Grand Tour Mean Distances", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
scatter_plot1 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="One-Day Races Total Distances")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```

```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

scatter_plot2 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="Stage Races Total Distances")
legend("topleft", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Total Distance - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Total Distance", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Total Distance - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Total Distance", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

scatter_plot3 <- plot(scatterTemp$Year, scatterTemp$`Total Distance`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Total Distance (km)', main="Grand Tour Races Total Distances")
legend("right", legend=c('Men', 'Women'), pch = 19, col=colors)
```


################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE NUMBER OF TEAMS BAR CHARTS
################################################################################

First, the mean number of teams for each year for the women's races is calculated, with the data values being entered manually, and with the data being split between the different possible categories of races
```{r}
wMeanTeams_df = data.frame(Year=c(2016:2024))
wMeanTeams_df$TeamStage = c(0)
wMeanTeams_df$TeamGT = c(0)
wMeanTeams_df$TeamOD = c(0)

for(i in 0:8) {
  wMeanTeams_df$TeamStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Num Teams - WOMEN`)
  wMeanTeams_df$TeamGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Num Teams - WOMEN`)
  if(i >= 7) {
    wMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Num Teams - WOMEN`)
  } else {
    wMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Num Teams - WOMEN`)
  }
}
```

Next, we find the same values for the men's races:
```{r}
mMeanTeams_df = data.frame(Year=c(2016:2024))
mMeanTeams_df$TeamStage = c(0)
mMeanTeams_df$TeamGT = c(0)
mMeanTeams_df$TeamOD = c(0)

for(i in 0:8) {
  mMeanTeams_df$TeamStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Num Teams - MEN`)
  mMeanTeams_df$TeamGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Num Teams - MEN`)
  mMeanTeams_df$TeamOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Num Teams - MEN`)
}
```

We then use this information to construct a sequence of bar charts:
```{r}
teams_matrix <- matrix(c(wMeanTeams_df$TeamOD, mMeanTeams_df$TeamOD), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's One-Day Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


teams_matrix <- matrix(c(wMeanTeams_df$TeamStage, mMeanTeams_df$TeamStage), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's Stage Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


teams_matrix <- matrix(c(wMeanTeams_df$TeamGT, mMeanTeams_df$TeamGT), nrow = 2, byrow = TRUE)
rownames(teams_matrix) <- c("Womens", "Mens")
colnames(teams_matrix) <- wMeanTeams_df$Year
barplot(teams_matrix, xlab="Year", ylab="Mean Number of Teams", main="Men and Women's Grand Tour Mean Number of Teams", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="One-Day Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="Stage Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Num Teams - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Num Teams", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Num Teams - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Num Teams", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Num Teams`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Number of Teams', main="Grand Tour Races Number of Teams")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```

################################################################################
FOLLOWING CODE IS USED FOR GENERATING THE PLOTS FOR FINAL WINNER TIMES
################################################################################

```{r}
wMeanTime_df = data.frame(Year=c(2016:2024))
wMeanTime_df$TimeStage = c(0)
wMeanTime_df$TimeGT = c(0)
wMeanTime_df$TimeOD = c(0)

for(i in 0:8) {
  wMeanTime_df$TimeStage[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Stage',]$`Final Winner Time - WOMEN`)
  wMeanTime_df$TimeGT[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'Grand Tour',]$`Final Winner Time - WOMEN`)
  if(i >= 7) {
    wMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi' | races$Race == 'La Vuelta de Espana'),]$`Final Winner Time - WOMEN`)
  } else {
    wMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'One-Day' | races$Race == 'Tour of Guangxi'),]$`Final Winner Time - WOMEN`)
  }
}
```

```{r}
mMeanTime_df = data.frame(Year=c(2016:2024))
mMeanTime_df$TimeStage = c(0)
mMeanTime_df$TimeGT = c(0)
mMeanTime_df$TimeOD = c(0)

for(i in 0:8) {
  mMeanTime_df$TimeStage[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Stage' | races$Race == 'Tour of Guangxi'),]$`Final Winner Time - MEN`)
  mMeanTime_df$TimeGT[9-i] <- mean(races[races$Year == (2024-i) & (races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana'),]$`Final Winner Time - MEN`)
  mMeanTime_df$TimeOD[9-i] <- mean(races[races$Year == (2024-i) & races$Type == 'One-Day',]$`Final Winner Time - MEN`)
}
```

```{r}
time_matrix <- matrix(c(wMeanTime_df$TimeOD, mMeanTime_df$TimeOD), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's One-Day Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


time_matrix <- matrix(c(wMeanTime_df$TimeStage, mMeanTime_df$TimeStage), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's Stage Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))


time_matrix <- matrix(c(wMeanTime_df$TimeGT, mMeanTime_df$TimeGT), nrow = 2, byrow = TRUE)
rownames(time_matrix) <- c("Womens", "Mens")
colnames(time_matrix) <- wMeanTime_df$Year
barplot(time_matrix, xlab="Year", ylab="Mean Final Winner Time in Minutes", main="Men and Women's Grand Tour Mean Final Winner Times (Minutes)", beside=TRUE, legend=c('Womens', 'Mens'), args.legend=list(x='bottomright'), col = c("lightblue", "lightcoral"))
```
```{r}
scatterTemp <- races[races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day", c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'One-Day', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

colors <- c("lightcoral", "lightblue") 
plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="One-Day Races Final Winner Times")
legend("bottomright", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Stage', c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'Stage' | races$Race == 'Tour of Guangxi', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="Stage Races Final Winner Times")
legend("topleft", legend=c('Men', 'Women'), pch = 19, col=colors)
```
```{r}
scatterTemp <- races[races$Type == 'Grand Tour', c('Year', 'Final Winner Time - WOMEN')]
scatterTemp$Gender <- 'Women'
colnames(scatterTemp) <- c("Year", "Final Winner Time", "Gender")

temp <- races[races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana', c('Year', 'Final Winner Time - MEN')]
temp$Gender <- 'Men'
colnames(temp) <- c("Year", "Final Winner Time", "Gender")

scatterTemp <- rbind(scatterTemp, temp)

plot(scatterTemp$Year, scatterTemp$`Final Winner Time`, col=colors[factor(scatterTemp$Gender)], pch = 19, xlab='Year', ylab='Final Winner Time', main="Grand Tour Races Final Winner Times")
legend("right", legend=c('Men', 'Women'), pch = 19, col=colors)
```

################################################################################
 FOLLOWING CODE IS USED FOR GENERATING THE BAR PLOTS FOR THE INDIVIDUAL AND 
 OVERALL PRIZE MONEY
################################################################################
 
First, we collect the data for the individual and overall prize money for the women's races, with these being separated by category:
```{r}
 individualWomenPrizeMoney2024Stage <- c(12000, 12030, 16000)
 individualWomenPrizeMoney2024GrandTour <- c(50000, 50000)
 individualWomenPrizeMoney2024OneDay <- c(16000, 16000, 2256, 16000, 20000, 21000, 16000, 10000, 12500)
 
 overallWomenPrizeMoney2024Stage <- c(100000, 100000, 130000)
 overallWomenPrizeMoney2024GrandTour <- c(250000, 250000)
 overallWomenPrizeMoney2024OneDay <- c(43000, 40000, 40000, 50000, 34000, 20000)
```
 
 
 We then collect the same data for the men's races:
```{r}
 individualMenPrizeMoney2024Stage <- c(12000, 35500, 12000, 16000, 12300, 20000)
 individualMenPrizeMoney2024GrandTour <- c(150000, 265000, 500000)
 individualMenPrizeMoney2024OneDay <- c(16000, 16000, 16000, 16000, 16000, 20000, 30000, 16000, 16000, 20000)
 
 overallMenPrizeMoney2024Stage <- c(100000, 300000, 99000, 130000, 103700, 128000)
 overallMenPrizeMoney2024GrandTour <- c(1119960, 1600000, 2573202)
 overallMenPrizeMoney2024OneDay <- c(43000, 40000, 40000, 40000, 40000, 50000, 91000, 34000, 40000, 50000)
```

A barplot is then constructed for the individual and overall in each category:
```{r}
ind_mean_tableOD <- c(mean(individualWomenPrizeMoney2024OneDay), mean(individualMenPrizeMoney2024OneDay))
barplot(ind_mean_tableOD, xlab="", ylab="Mean Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024OneDay) + 5000), col=c("lightblue", "lightcoral"), main="One-Day Races Mean Individual Winner Prize Money", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

ind_mean_tableS <- c(mean(individualWomenPrizeMoney2024Stage), mean(individualMenPrizeMoney2024Stage))
barplot(ind_mean_tableS, xlab="", ylab="Mean Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024Stage) + 5000), col=c("lightblue", "lightcoral"), main="Stage Races Mean Individual Winner Prize Money", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

ind_mean_tableGT <- c(mean(individualWomenPrizeMoney2024GrandTour), mean(individualMenPrizeMoney2024GrandTour))
barplot(ind_mean_tableGT, xlab="", ylab="Mean Individual Winner Prize Money (Euros)", ylim=c(0, mean(individualMenPrizeMoney2024GrandTour) + 5000), col=c("lightblue", "lightcoral"), main="Grand Tour Races Mean Individual Winner Prize Money", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableOD <- c(mean(overallWomenPrizeMoney2024OneDay), mean(overallMenPrizeMoney2024OneDay))
barplot(overall_mean_tableOD, xlab="", ylab="Mean Overall Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024OneDay) + 5000), col=c("lightblue", "lightcoral"), main="One-Day Races Mean Overall Prize Pool", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableS <- c(mean(overallWomenPrizeMoney2024Stage), mean(overallMenPrizeMoney2024Stage))
barplot(overall_mean_tableS, xlab="", ylab="Mean Overall Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024Stage) + 5000), col=c("lightblue", "lightcoral"), main="Stage Races Mean Overall Prize Pool", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))

overall_mean_tableGT <- c(mean(overallWomenPrizeMoney2024GrandTour), mean(overallMenPrizeMoney2024GrandTour))
barplot(overall_mean_tableGT, xlab="", ylab="Mean Overall Prize Pool (Euros)", ylim=c(0, mean(overallMenPrizeMoney2024GrandTour) + 5000), col=c("lightblue", "lightcoral"), main="Grand Tour Races Mean Overall Prize Pool", legend=c("Women's", "Men's"), args.legend=list(x="bottomright", col=c("lightblue", "lightcoral")))
```


################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING DISTANCES IN 2024
################################################################################

Testing the difference in total distances for one day races in 2024. Our alternative hypothesis is that the women's races are significantly lower in distance than the men's races

```{r}
fDistOD24 <- races[(races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day") & races$Year == 2024, c('Total Distance - WOMEN')]

mDistOD24 <- races[(races$Type == 'One-Day') & races$Year == 2024, c('Total Distance - MEN')]

t.test(fDistOD24[[1]], mDistOD24[[1]], alternative = 'less')
```

Testing the difference in total distances for stage races in 2024. Our alternative hypothesis is that the women's races are significantly lower in distance than the men's races

```{r}
fDistStage24 <- races[races$Type == 'Stage' & races$Year == 2024, c('Total Distance - WOMEN')]

mDistStage24 <- races[(races$Type == 'Stage' | races$Race == 'Tour of Guangxi') & (races$Year == 2024), c('Total Distance - MEN')]

t.test(fDistStage24[[1]], mDistStage24[[1]], alternative = 'less')
```

Testing the difference in total distances for grand tour in 2024. Our alternative hypothesis is that the women's races are significantly lower in distance than the men's races

```{r}
fDistGT24 <- races[races$Type == 'Grand Tour' & races$Year == 2024, c('Total Distance - WOMEN')]

mDistGT24 <- races[(races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana') & races$Year == 2024, c('Total Distance - MEN')]

t.test(fDistGT24[[1]], mDistGT24[[1]], alternative = 'less')
```

################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING DISTANCES FROM 2016-2024
################################################################################

```{r}
temp = races[c('Race', 'Type', 'Total Distance - WOMEN', 'Total Distance - MEN', 'Year')]

# had to pull out tidyverse for this one
dist_long <- temp %>% pivot_longer(
  cols = c('Total Distance - WOMEN', 'Total Distance - MEN'),
  names_to = "Gender",
  values_to = "Distance"
  ) %>%
  mutate(
    Gender = ifelse(Gender == "Total Distance - MEN", "Men", "Women")
  ) 

one_day <- dist_long %>% filter(Type == "One-Day" | (Race=='Tour of Guangxi' & Gender == 'Women') | (Race=='La Vuelta de Espana' & Gender == 'Women' & Year < 2018))
stage <- dist_long %>% filter(Type == "Stage" | (Race=='Tour of Guangxi' & Gender == 'Men'))
grand_tour <- dist_long %>% filter(Type == "Grand Tour" | (Race=='La Vuelta de Espana' & Gender == 'Men'))
```

The first model tests whether there is a consistent difference between men's and women's one day race distances across all years. A significant effect of gender in this model indicates a consistent gender gap.

```{r}
model_one_day <- lm(Distance ~ Gender + Year, data = one_day)
summary(model_one_day)
```

The second model adds an interaction term between gender and year to test whether the gender gap for distances has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_one_day <- lm(Distance ~ Gender * Year, data = one_day)
summary(interaction_model_one_day)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.

```{r}
anova(model_one_day, interaction_model_one_day)
```
There is a statistically significant difference in one day race distances between men and women across all years (p < 2e-16). However, the interaction between gender and year is not significant, suggesting that the gender gap has remained relatively stable over time.


#######

The first model tests whether there is a consistent difference between men's and women's stage race distances across all years. A significant effect of gender in this model indicates a consistent gender gap.
```{r}
model_stage <- lm(Distance ~ Gender + Year, data = stage)
summary(model_stage)
```

The second model adds an interaction term between gender and year to test whether the gender gap for distances has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_stage <- lm(Distance ~ Gender * Year, data = stage)
summary(interaction_model_stage)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.
```{r}
anova(model_stage, interaction_model_stage)
```
There is a statistically significant difference in one day race distances between men and women across all years (p < 2e-16). However, the interaction between gender and year is not significant, suggesting that the gender gap has remained relatively stable over time.


####

The first model tests whether there is a consistent difference between men's and women's grand tour race distances across all years. A significant effect of gender in this model indicates a consistent gender gap.
```{r}
model_grand_tour <- lm(Distance ~ Gender + Year, data = grand_tour)
summary(model_grand_tour)
```

The second model adds an interaction term between gender and year to test whether the gender gap for distances has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_grand_tour <- lm(Distance ~ Gender * Year, data = grand_tour)
summary(interaction_model_grand_tour)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.
```{r}
anova(model_grand_tour, interaction_model_grand_tour)
```

There is a statistically significant difference in one day race distances between men and women across all years (p < 2e-16). However, the interaction between gender and year is not significant, suggesting that the gender gap has remained relatively stable over time.

################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING THE NUMBER OF TEAMS IN 2024  
################################################################################

Testing the difference in team amounts for one day races in 2024. Our alternative hypothesis is that the women's races are significantly lower in team amounts than the men's races

```{r}
fTeamsOD24 <- races[(races$Type == 'One-Day' | races$Type == "Grand Tour / women's one-day" | races$Type == "Men's stage, women's one-day") & races$Year == 2024, c('Num Teams - WOMEN')]

mTeamsOD24 <- races[(races$Type == 'One-Day') & races$Year == 2024, c('Num Teams - WOMEN')]

t.test(fTeamsOD24[[1]], mTeamsOD24[[1]], alternative = 'less')
```

Testing the difference in team amounts for stage races in 2024. Our alternative hypothesis is that the women's races are significantly lower in team amounts than the men's races
```{r}
fTeamsStage24 <- races[races$Type == 'Stage' & races$Year == 2024, c('Num Teams - WOMEN')]

mTeamsStage24 <- races[(races$Type == 'Stage' | races$Race == 'Tour of Guangxi') & (races$Year == 2024), c('Num Teams - WOMEN')]

t.test(fTeamsStage24[[1]], mTeamsStage24[[1]], alternative = 'less')
```

Testing the difference in team amounts for grand tour races in 2024. Our alternative hypothesis is that the women's races are significantly lower in team amounts than the men's races
```{r}
fTeamsGT24 <- races[races$Type == 'Grand Tour' & races$Year == 2024, c('Num Teams - WOMEN')]

mTeamsGT24 <- races[(races$Type == 'Grand Tour' | races$Race == 'La Vuelta de Espana') & races$Year == 2024, c('Num Teams - WOMEN')]

t.test(fTeamsGT24[[1]], mTeamsGT24[[1]], alternative = 'less')
```

################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING TEAM AMOUNTS FROM 2016-2024
################################################################################

```{r}
temp = races[c('Race', 'Type', 'Num Teams - WOMEN', 'Num Teams - MEN', 'Year')]

# had to pull out tidyverse for this one
team_long <- temp %>% pivot_longer(
  cols = c('Num Teams - WOMEN', 'Num Teams - MEN'),
  names_to = "Gender",
  values_to = "Teams"
  ) %>%
  mutate(
    Gender = ifelse(Gender == "Num Teams - MEN", "Men", "Women")
  ) 

one_day <- team_long %>% filter(Type == "One-Day" | (Race=='Tour of Guangxi' & Gender == 'Women') | (Race=='La Vuelta de Espana' & Gender == 'Women' & Year < 2018))
stage <- team_long %>% filter(Type == "Stage" | (Race=='Tour of Guangxi' & Gender == 'Men'))
grand_tour <- team_long %>% filter(Type == "Grand Tour" | (Race=='La Vuelta de Espana' & Gender == 'Men'))
```

The first model tests whether there is a consistent difference between men's and women's one day race team amounts across all years. A significant effect of gender in this model indicates a consistent gender gap.

```{r}
model_one_day <- lm(Teams ~ Gender + Year, data = one_day)
summary(model_one_day)
```

The second model adds an interaction term between gender and year to test whether the gender gap for team amounts has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_one_day <- lm(Teams ~ Gender * Year, data = one_day)
summary(interaction_model_one_day)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.

```{r}
anova(model_one_day, interaction_model_one_day)
```

There is a statistically significant difference in one day race team amounts between men and women across all years (p < 0.00629). However, the interaction between gender and year is not significant, suggesting that the gender gap for team amounts has remained relatively stable over time.

#
Although the 2024 data alone does not show a statistically significant gender difference in the number of teams (t-test, p = 0.3635), a linear model across all years reveals a significant gender effect. This suggests that while the gender gap may not appear in 2024 specifically, it is consistent and detectable over the full time span
#


#######

The first model tests whether there is a consistent difference between men's and women's stage race team amounts across all years. A significant effect of gender in this model indicates a consistent gender gap.

```{r}
model_stage <- lm(Teams ~ Gender + Year, data = stage)
summary(model_stage)
```

The second model adds an interaction term between gender and year to test whether the gender gap for team amounts has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_stage <- lm(Teams ~ Gender * Year, data = stage)
summary(interaction_model_stage)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.

```{r}
anova(model_one_day, interaction_model_one_day)
```
There is a statistically significant difference in stage race team amounts between men and women across all years (p < 8.85e-07). However, the interaction between gender and year is not significant, suggesting that the gender gap for team amounts has remained relatively stable over time.


#######

The first model tests whether there is a consistent difference between men's and women's grand tour race team amounts across all years. A significant effect of gender in this model indicates a consistent gender gap.

```{r}
model_grand_tour <- lm(Teams ~ Gender + Year, data = grand_tour)
summary(model_grand_tour)
```
The second model adds an interaction term between gender and year to test whether the gender gap for team amounts has changed over time (i.e., whether it has increased or decreased). If the interaction is not significant, this suggests that the more complex model does not provide a substantially better fit than the simpler additive model.
```{r}
interaction_model_grand_tour <- lm(Teams ~ Gender * Year, data = grand_tour)
summary(interaction_model_grand_tour)
```

Comparing the two models using ANOVA, we found a high p-value for the interaction term, indicating that it is not statistically significant.

```{r}
anova(model_one_day, interaction_model_one_day)
```
There is not a statistically significant difference in grand tour race team amounts between men and women across all years (p < 0.991). The interaction between gender and year is not significant either, suggesting that there has not been a consistent difference in the amount of teams for men and women's grand tour races.


################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING WITH INDIVIDUAL PRIZE MONEY
################################################################################

Testing the difference in individual prize money for one day races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(individualWomenPrizeMoney2024OneDay, individualMenPrizeMoney2024OneDay, alternative='less')
```

Testing the difference in individual prize money for stage races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(individualWomenPrizeMoney2024Stage, individualMenPrizeMoney2024Stage, alternative='less')
```

Testing the difference in individual prize money for grand tour races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(individualWomenPrizeMoney2024GrandTour, individualMenPrizeMoney2024GrandTour, alternative='less')
```

################################################################################
 FOLLOWING CODE IS USED FOR HYPOTHESIS TESTING WITH OVERALL PRIZE MONEY
################################################################################

Testing the difference in overall prize money for one day races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(overallWomenPrizeMoney2024OneDay, overallMenPrizeMoney2024OneDay, alternative='less')
```

Testing the difference in overall prize money for stage races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(overallWomenPrizeMoney2024Stage, overallMenPrizeMoney2024Stage, alternative='less')
```

Testing the difference in overall prize money for grand tour races in 2024. Our alternative hypothesis is that the women's races are significantly lower in prize money amounts than the men's races
```{r}
t.test(overallWomenPrizeMoney2024GrandTour, overallMenPrizeMoney2024GrandTour, alternative='less')
```




